---
title: "EDAProject"
author: "KenDopp"
date: "Saturday, July 11, 2015"
output: html_document
---
#Citation   
  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

  Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
                [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib


#load in the data
```{r}
getwd()
setwd("C:/Users/Kenneth/Desktop/UdacityFC/Intro to R/IntrotoRFinalProject/")
list.files()
wd<- read.csv('wineQualityReds.csv')
```

#do some exploring, my first step is to see how all of the 12 variables relate to quality and to spot out any patterns
```{r}
names(wd)
library(ggplot2)


ggplot(aes(x = alcohol, y = quality), data = wd) + geom_point(position = 'jitter')
#seems to be somewhat of a correlation between how good the wines are and how much alcohol they have


ggplot(aes(x = free.sulfur.dioxide, y = quality), data = wd) + geom_point(position = 'jitter')
#somewhat of a normal distribution between sulfur dioxide and quality, with both the worst and best wines 
#tho, the distribution is somewhat skewed so to improve wine quality a quick check could just to check the sulfure dioxide as if it's above 30 it's basically guaranteed that it is at least an average (i.e. not horrible) wine



ggplot(aes(x = chlorides, y = quality), data = wd) + geom_point(position = 'jitter')
#somewhat similar to sulfur dioxide, high concentrations lead to an average quality


ggplot(aes(x = X, y = quality), data = wd) + geom_point(position = 'jitter')
#I cant make out any patterns here, seems to be totally unrelated to quality

ggplot(aes(x = fixed.acidity, y = quality), data = wd) + geom_point(position = 'jitter')
#same pattern, exteme values produce average wine

ggplot(aes(x = volatile.acidity, y = quality), data = wd) + geom_point(position = 'jitter')
#weak negative correlation between volatile acidity and quality, all over 1 are less than average

ggplot(aes(x = citric.acid, y = quality), data = wd) + geom_point(position = 'jitter')
#kinda similar pattern to other variables like X or fixed acidity, only here great wines can have a high citric 
#acid amount as well as average ones, thus having a high citric acid amount leads to at least decent quality
#tho it's not really correlated in the sense that more leads to high quality, that only works to a til it reaches a quality of about 5

ggplot(aes(x = residual.sugar, y = quality), data = wd) + geom_point(position = 'jitter')
#looks basically the same as chlorides, same pattern of more to get average quality and sort of a normal distribution,
#so lots of sugar is a good indication that it's a decent wine, and low sugar increases the probability of a great
#or horrible quality wine

ggplot(aes(x = total.sulfur.dioxide, y = quality), data = wd) + geom_point(position = 'jitter')
#normal distribution pattern, also very similar to free sulfur dioxide, two outliers here that seem to suggest
#exteme values don't diminish quality, but the general pattern is clearly more leads to average quality

ggplot(aes(x = density, y = quality), data = wd) + geom_point(position = 'jitter')
#no real correlation but extremely low or high densities indicate average quality
#on second thought, there appears to be a weak negative correlation between density and quality, less dense is better

ggplot(aes(x = pH, y = quality), data = wd) + geom_point(position = 'jitter')
#seems like no correlation whatsoever

ggplot(aes(x = sulphates, y = quality), data = subset(wd, wd$sulphates <= quantile(wd$sulphates, 0.99))) + geom_point(alpha = 1/2, position = 'jitter') 
#noticeable correlation with extreme values indicating average quality but beneath about 1/ cutting out extreme
#values appears to be a decent correlation
#seems highly correlated up until about 0.75, then above this value means that the wine is not horrible but it's basically uncorrelated as long as it's above 4.5
#make graph of filtered sulphates to see if hypothesized relationship holds, also look at correlation


#using ggpairs to get an overview

library(GGally)
library(scales)
library(memisc)
ggpairs(wd)
#key takeaway is high correlation with alcohol(highest), strong negative with citric.acid, and some others decently strong

#do some qplots to see the numbers at different ratings
qplot(data = wd, x = quality, binwidth = 0.5)
#normal, slightly skewed left curve, few under 4 or above 8

```
#So I noticed a few things here, there are basically 2 categories, ones like Alchohol seems to have a fairly 
#straightforward linear relationship. Others like residual sugar have a sort of normal distribution where extreme positive values are an indicator of average quality (so could be a useful test to make sure your wine is at least decent), however extreme values filter out both low and high quality wines. Sulphates is a sort of other category

#exhibiting both a linear relationship and the extreme values meaning high quality

#qualities that appear to have no relation to quality:
X, pH
#qualities that seem to have somewhat of a direct correlation
Alcohol
volatile.acidity
residual.sugar
density
sulphates
#super high values means mediocre, jacking up these valus should guarantee a saleable wine
free.sulfur.dioxide
chlorides
fixed.acidity
citric.acid
total.sulfur.dioxide

#linear model of quality
```{r}
m1 <- lm(I(quality) ~ I(citric.acid), data = wd)
m2 <- update(m1, ~ . + sulphates)
mtable(m1)


```

